---
title: "Lab 3"
author: "DATA1220-55"
date: '`r Sys.Date()`'
format: 
  pdf:
    engine: knitr
    execute:
      echo: true
      eval: false
      message: false
      warning: false
    self-contained: true
---

***Research Question: Do babies born to mothers who smoke have lower birth weights than babies born to non-smokers?***

1.  Start a new project in RStudio.

2.  Start a new R script and save it as `lab03.R`.

3.  You will need to use the `dplyr` library for this lab. Load it now using the `library()` function.

4.  Download the Excel spreadsheet `births14.xlsx`.

    a.  Go to the URL <https://github.com/sarah-grabinski/data1220-55_fall2024/raw/refs/heads/main/slides/lab01/births14.xlsx> and manually download the file.

    b.  Modify the code below to download the file directly to your project folder using the URL <https://github.com/sarah-grabinski/data1220-55_fall2024/raw/refs/heads/main/slides/lab01/births14.xlsx>.

    ```{r}
    download.file("http://www.website.com/excel_spreadsheet.xlsx", 
                  destfile = "excel_spreadsheet.xlsx", 
                  mode = "wb")
    ```

5.  Use the "Import Dataset" tool in RStudio to load the file `births14.xlsx` into RStudio as the dataframe `births14`.

6.  Modify the function below to inspect your data using the `summary()` function.

    a.  What is the value of the the 3rd quartile of the variable `weight`?
    b.  Compare the mean and median of the variable `weight`. Is it reasonable to assume normality here?

    ```{r}
    summary(dataframe)
    ```

7.  Generate summary statistics $\bar{x}$ for the mean birth weight `weight` of babies by smoking habit `habit` using the `summarize()` function from the `dplyr` library and the `mean()` function. You will also need to estimate the standard deviation $s$ with the `sd()` function and get the sample size $n$ with the `n()` function for each group. Modify the example below.

    ```{r}
    dataframe |>
      summarize(newcolumn1 = function1(), 
                newcolumn2 = function2(variable1), 
                newcolumn3 = function3(variable1), 
                .by = "variable2")
    ```

8.  Calculate a point estimate for the difference $\bar{x}_1-\bar{x}_2$ between the mean birth weight of babies born to smoking moms and the mean birth weight of babies born to non-smoking moms. Store the result as a variable called `mean_diff`.

    ```{r}
    new_variable <- 1 + 2
    ```

9.  Calculate the standard error $\text{SE}$ for the difference between the 2 mean birth weights $\bar{x}_1-\bar{x}_2$. Store the result as a variable called `se_diff`.

    $$
     \text{SE}=\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}
    $$

    ```{r}
    new_variable <- sqrt((s1^2)/n1 + (s2^2)/n2)
    ```

10. Construct a 95% confidence interval using the Student's $t$ distribution for the difference in average birth weights by smoking habit.

    a.  Calculate the degrees of freedom for this $t$ distribution and store it as the variable `dof`.

        $$
         \text{degrees of freedom} = \text{min}(n_1, n_2)-1
        $$

    b.  Find the value of alpha $\alpha$ for a 95% confidence interval and store it as the variable `alpha`.

        $$
         \alpha=1-\text{confidence}
        $$

    c. Find the critical value $T^*$ for a 95% confidence interval using the `qt(probability, df)` function. Use the probability $\alpha/2$ or $1-\alpha/2$ and the degrees of freedom `dof` that you just calculated. Store it as the variable `t_star`.

    d. Calculate the upper and lower bounds of your confidence interval as $\text{point estimate} + T^* \times \text{SE}$ and $\text{point estimate} - T^* \times \text{SE}$.
    
    e. Can you interpret your confidence interval in a complete sentence?

    f. Based on the boundaries of your confidence interval, can you make any early guesses about whether your hypothesis test will be significant?


11. Perform a 2-sample $t$-test to test the hypothesis that there is a difference in average birth weights between babies born to smoking and non-smoking mothers.

      $$
        H_0 \colon \mu_{\text{smoker}}-\mu_{\text{nonsmoker}}=0
      $$
      $$
        H_A \colon \mu_{\text{smoker}}-\mu_{\text{nonsmoker}}<0
      $$

    a. Calculate the test statistic $T$ for your observed difference under the null hypothesis that there is no difference $H_0 \colon \mu_{\text{smoker}}-\mu_{\text{nonsmoker}}=0$. Save the result as the variable `test_statistic`.
    
        $$
          `T = \frac{\bar{x}_1-\bar{x}_2}{SE}
        $$
    
    b. Get the p-value for the one-sided hypothesis test (left-tailed) from your `test_statistic` $T$ using the `pt()` function and your degrees of freedom `dof`. Depending on the value of your test statistic $T$, you may need to change the `lower.tail` parameter to false `F` to get the proper probability.
    
        ```{r}

        pt(statistic, df, lower.tail = T)

        ```
    
    c. Using the significance level `alpha` or $\alpha$ that you calculated previously, would you reject the null hypothesis? Why or why not?
        