---
title: "Untitled"
format: revealjs
editor: visual
---

## Data for Today

-   Source: [National Health and Nutrition Examination Survey](https://www.cdc.gov/nchs/nhanes/index.htm) (NHANES), a program of studies designed to assess the health and nutritional status of adults and children in the United States

-   Survey combines data from interviews (demographic, socioeconomic, dietary, and health-related questions), physical examinations, and laboratory tests.

-   This data set contains 5000 observations from the 2009-2010 survey and 5000 observations from the 2011-2012 survey.

## Load NHANES data from `NHANES` package

```{r}
# this data comes from the NHANES package
# which you'll need to install before using
library(NHANES)
# this function pushes the data out of the package
# and into your global environment
force(NHANES)

NHANES
```

## Data Dictionary

| Variable Name  | Description                                                                 | Variable Type         |
|----------------|-----------------------------------------------------------------------------|-----------------------|
| `ID`           | a unique identifier for each of the obserations                             | categorical           |
| `Age`          | age of subject in years                                                     | numerical, discrete   |
| `AgeDecade`    | age range subject falls in                                                  | categorical, ordinal  |
| `BMI`          | body mass index                                                             | numerical, continuous |
| `Height`       | subject's height, in cm                                                     | numerical, continuous |
| `Testosterone` | testosterone levels from blood test                                         | numerical, continuous |
| `AlcoholDay`   | average number of alcoholic drinks consumed per day over the past 12 months | numerical, discrete   |

## Generating numerical summaries in R: `summary()` function from `base` R

```{r}
# pipe your data into your workflow
NHANES |>
  # select your variable(s) of interest
  select(Age) |>
  # run the summary function
  summary()
```

## Generating numerical summaries in R: `Hmisc::describe()`

```{r}
# pipe your data into your workflow
NHANES |>
  # select your variable(s) of interest
  select(Age) |>
  # run the describe() function from the Hmisc package
  Hmisc::describe()
```

```{r}
NHANES |>
  ggplot(aes(x = Weight)) +
  geom_density()

NHANES |>
  ggplot(aes(x = Height)) +
  geom_density()
```
