---
title: "Homework 4"
author: "YOURNAMEHERE"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    theme: journal
    self-contained: true
editor: visual
---

# Setup  {.unnumbered}

Load the packages you will need to complete this analysis in the code chunk below.

```{r}
library(readxl)
library(tidyverse)

theme_set(theme_bw())
```

Download the file `nhanes.xlsx` from Canvas. Load the data into R using the function `read_excel()` from the `readxl` package and save it as the dataframe `nhanes` in the code chunk below.

```{r}
path_to_folder <- getwd()
file_name <- 'nhanes.xlsx'
path_to_file <- paste(path_to_folder, 
                      file_name, 
                      sep = '/')
nhanes <- read_xlsx(path_to_file)
```

# Instructions  {.unnumbered}

-   Use complete sentences where indicated.

-   Save your file as `homework4_yournamehere.qmd` and renter it as an HTML file (`homework4_yournamehere.html`).

-   Submit both `homework4_yournamehere.qmd` and `homework4_yournamehere.html` to Canvas by November 8th, 2024 at 6:00pm.

# Problem 1
